\huge{ \emph{Supplement to} Algorithms For Shaping a Particle Swarm\\ With a Shared Control Input Using Boundary Interaction}

### 2016-09-10 RA-L/theory.tex
Changed some wrong notations
Full revision pass.

The region of integration $R$ is the polygon containing the swarm. If the force angle is $\beta$, the mean when the swarm is in the lower-left corner is:
\begin{align}\label{eq:meanInSquareWorkspaceLL}
\bar{x}(A,\beta) &= \frac{\int_0^{\sqrt{2} \sqrt{-A \tan (\beta )}} \left(\int_0^{\sqrt{2} \sqrt{-A \cot (\beta )}+x \cot (\beta )} x \, dy\right) \, dx}{A} \nonumber \\
	&=\frac{ \sqrt{2}}{3} \sqrt{A \tan (\beta )}\\
\bar{y}(A,\beta) &= \frac{\int_0^{\sqrt{2} \sqrt{-A \tan (\beta )}} \left(\int_0^{\sqrt{2} \sqrt{-A \cot (\beta )}+x \cot (\beta )} y \, dy\right) \, dx}{A} \nonumber\\
	&=\frac{\sqrt{2}}{3}  \sqrt{A \cot (\beta )}
\end{align}
The full equations are included in the appendix~\citep{2016arXiv160901830S}, and are summarized in Fig.~\ref{fig:SquareFill}. A few highlights are that the correlation is maximized $\pm$1/2 when the swarm is in a triangular shape. The covariance of a triangle is always $\pm(A/18)$. Variance is minimized in the direction of $\beta$ and maximized orthogonal to $\beta$ when the swarm is in a rectangular shape. The range of mean positions are maximized when $A$ is small.

@@

\subsection{Using Boundaries: Friction and Boundary Layers}\label{subsec:WallFriction}
Global inputs move a swarm uniformly.  
Controlling covariance requires breaking this uniform symmetry.  A swarm inside an axis-aligned rectangular workspace can reduce variance normal to a wall by simply pushing the swarm into the boundary. Directly controlling covariance by pushing the swarm into a boundary requires changes to the boundary.  An obstacle in the lower-right corner is enough to generate positive covariance, but generating both positive and negative covariance requires additional obstacles.  Requiring a special obstacle configuration also makes covariance control dependent on the local environment. 
  Instead of pushing our robots directly into a wall, this paper examines an oblique approach, by using boundaries that generate friction with the robots.  These frictional forces are  sufficient to break the symmetry caused by uniform inputs.  Robots touching a wall have a negative friction force that opposes movement along the boundary.  This  causes robots along the boundary to move more slowly than robots in free-space.

Let the control input be a vector force $\vec{F}$ with magnitude $F$ and orientation $\theta$ with respect to a line perpendicular to and into the nearest boundary. $N$ is the normal or perpendicular force between the robot and the boundary. The force of friction $F_f$ is nonzero if the robot is in contact with the boundary and  $|\theta| < \pi/2$. The resulting net force on the robot, $F_{\text{\emph{forward}}}$, is aligned with the wall and given by
\begin{align}
F_{\text{\emph{forward}}} &=  F \sin(\theta) - F_f  \nonumber \\
\text{where }  F_f &= \begin{cases}  \mu_f N, &  \mu_f N < F \sin(\theta)  \label{eq:frictionmodel}  \\
F \sin(\theta), & \text{else} \end{cases} \\ %\sign(F \sin(\theta) ) \cdot  \max(0, | F sin\theta |- |F_f|)
\text{and } N &= F \cos(\theta) \nonumber
\end{align}
 Fig.~\ref{fig:friction} shows the resultant forces on two robots when one is touching a wall. 
Though each receives the same inputs,  they experience different net forces.
  For ease of analysis, the following algorithms assume $\mu_f$ is infinite and robots touching the wall are prevented from sliding along the wall.
This means that if one robot is touching the wall and another robot is free, the touching robot will not move when the control input is parallel or into the wall. 
There are many alternate models of friction that also break control symmetry. Fig.~\ref{fig:friction}c shows fluid flow along a boundary.  Fluid in the free-flow region moves uniformly, but flow decreases to zero in the boundary layer.  
\begin{align}
u(y) = u_0 [1- \frac{(y-h)^2}{h^2} = u_0 \frac{y}{h} [2- \frac{y}{h}] \label{eq:boundarylayerflow}
\end{align}

The next section shows how a system with friction model \eqref{eq:frictionmodel} and two orthogonal walls can arbitrarily position two robots. 
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\columnwidth]{friction.pdf} 
\vspace{-1em}
\caption{(a,b) Wall friction reduces the force for going forward $F_{\text{\emph{forward}}}$ on a robot near a wall, but not for a free robot. (c) velocity of a fluid reduces to zero at the boundary.}
\label{fig:friction}
\end{center}
\end{figure}

@@

Our algorithm solves the positioning problem in two steps: 
First, $|\Delta r_x - \Delta e_x |$ is reduced to zero while  $\Delta r_y$ is kept constant in Alg.~\ref{alg:XControl}. 
Second $|\Delta r_y - \Delta e_y |$ is reduced to zero while  $\Delta r_x$ is kept constant.% in Alg.~\ref{alg:YControl}.

@@

\begin{algorithm}
\caption{WallFrictionArrange2Robots($s_1,s_2,e_1,e_2,L$)}\label{alg:PosControl2Robots}
\begin{algorithmic}[1]
\Require starting $(s_1,s_2)$ and ending $(e_1,e_2)$ positions of  two robots. 
$(0,0)$ is bottom corner, $s_1$ is rightmost robot, 
 $L$ is length of the walls. 
 Current position of the robots are $(r_1,r_2)$.

\State ($r_1,r_2$) = GenerateDesired$x$-spacing($s_1,s_2,e_1,e_2,L$)
\State GenerateDesired$y$-spacing($r_1,r_2,e_1,e_2,L$)

@@

\begin{algorithm}
\caption{GenerateDesired$x$-spacing($s_1,s_2,e_1,e_2,L$)}\label{alg:XControl}
\begin{algorithmic}[1]
\Require Knowledge of starting $(s_1,s_2)$ and ending $(e_1,e_2)$ positions of  two robots. 
$(0,0)$ is bottom corner, $s_1$ is topmost robot, 
 $L$ is length of the walls. Current robot positions are $(r_1,r_2)$.
\Ensure   $ r_{1y} - r_{2y}  \equiv s_{1y} - s_{2y} $   %$\Delta y(t) \equiv \Delta y(0)$ 
\State $\epsilon \gets $ small number
\State $ \Delta s_x  \gets s_{1x} - s_{2x} $
\State $ \Delta e_x \gets e_{1x} - e_{2x} $
\State $ r_1 \gets s_1$, $ r_2 \gets s_2$
\If {$\Delta e_x < 0 $ }
\State $ m \gets ( L-\epsilon-\max( r_{1x},r_{2x}) ,0)   $ \Comment{Move to right wall}
\Else 
\State  $ m \gets ( \epsilon-\min( r_{1x},r_{2x}),0 )    $ \Comment{Move to left wall}
\EndIf
\State $m  \gets  m + (0, -\min( r_{1y},r_{2y} ))$ \Comment{Move to bottom}
\State $ r_1 \gets r_1+m$, $ r_2 \gets r_2+m$ \Comment{Apply move}
\If {$\Delta e_x - (r_{1x} - r_{2x} ) > 0 $}
\State $ m \gets (\min(|\Delta e_x - \Delta s_x |, L- r_{1x}), 0)$  \Comment{Move right}
\Else
\State $ m \gets (-\min(|\Delta e_x - \Delta s_x |, r_{1x}), 0)$\Comment{Move left}
\EndIf 
\State $m  \gets  m + (0, \epsilon)$ \Comment{Move up}
\State $ r_1 \gets r_1+m$, $ r_2 \gets r_2+m$ \Comment{Apply move}
\State $\Delta r_x = r_{1x} - r_{2x}$
\If {$\Delta r_x \equiv \Delta e_x$} 
%\State   $ m \gets (e_{1x}-r_{1x}, e_{1y}-r_{1y})$
%\State $ r_1 \gets r_1+m$, $ r_2 \gets r_2+m$ \Comment{Apply move}
\State  \Return $(r_1,r_2)$
\Else   
\State \Return GenerateDesired$x$-spacing($r_1,r_2,e_1,e_2,L$)
\EndIf
\end{algorithmic}
\end{algorithm}

@@

\begin{figure}
\begin{center}
	\includegraphics[width=.9\columnwidth]{driftmove.pdf}
\end{center}
\vspace{-1em}
\caption{\label{fig:driftmove}
A  $\operatorname{DriftMove}(\alpha, \beta, \epsilon)$ to the right repeats a triangular movement sequence $\{ (\beta/2,-\epsilon),(\beta/2,\epsilon),(-\alpha,0)\}$. Robot $A$ touching a top wall moves right $\beta$ units, while robots not touching the top move right $\beta-\alpha$.
}
\end{figure}
\begin{figure}
\begin{center}
	\includegraphics[width=1.0\columnwidth]{multirobotSliderHardwareComp.pdf}
\end{center}
\vspace{-1em}
\caption{\label{fig:construction2d}
Illustration of Alg.\ \ref{alg:PosControlNRobots}, $n$ robot position control  using wall friction.
}
\end{figure}

Assume an open workspace with four axis-aligned walls with infinite friction.
The axis-aligned build zone of dimension $(w_b, h_b)$ containing the final configuration of $n$ robots must be disjoint from the axis-aligned staging zone of dimension $(w_s, h_s)$  containing the starting configuration of $n$ robots. Without loss of generality, assume the build zone  is above the staging zone. 
Furthermore, there must be at least $\epsilon$ space above the build zone, $\epsilon$ below the staging zone, and $\epsilon + 2r$ to the left of the build and staging zone, where $r$ is the radius of a robot.  The minimum workspace is then $(\epsilon + 2r + \max(w_f,w_s), 2\epsilon + h_s,h_f)$.

@@

Let $(0,0)$ be the lower left corner of the workspace, $p_k$ the $x,y$ position of the $k$th robot, and $f_k$ the final $x,y$ position of the $k$th robot. Label the robots in the staging zone from left-to-right and top-to-bottom, and the $f_k$ configurations right-to-left and top-to-bottom as shown in Fig.~\ref{fig:construction2d}.

\begin{algorithm}
\caption{PositionControl$n$RobotsUsingWallFriction($k$)}\label{alg:PosControlNRobots}
\begin{algorithmic}[1]
\State move( $-\epsilon, r-p_{k,y}$) % move  away from right wall and down till robot k touches bottom

\While{ $p_{k,x} > r$} 
\State $\operatorname{DriftMove}(\epsilon, \min(p_{k,x} - r,\epsilon), \epsilon)$ left   %drift move left until kth robot touches left wall
\EndWhile

\State $m \gets \operatorname{ceil}(\frac{f_{k,y}-r}{\epsilon})$
\State $\beta \gets \frac{f_{k,y}-r}{m}$
\State $\alpha \gets \beta - \frac{r - p_{k,y}-\epsilon}{m}$
\For{ $m$ iterations}
\State $\operatorname{DriftMove}(\alpha, \beta, \epsilon)$ up   %move kth robot to f_{ky} and leave the rest in position.
\EndFor

\State move($r+\epsilon-f_{k,x}, 0$)  % move the group to the left until k is in the correct relative x position
\State move($f_{k,x}-r, 0$)

@@

Alg. \ref{alg:PosControlNRobots} proceeds as follows:  
First, the robots are moved left away from the right wall, and down so robot $k$ touches the bottom wall.
Second, a set of $\operatorname{DriftMove}$s are executed that move robot $k$ to the left wall with no net movement of the other robots.
Third, a set of $\operatorname{DriftMove}$s are executed that  move robot $k$ to its target height and return the other robots to their initial heights. 
Fourth, all robots except robot $k$ are pushed left until robot $k$ is in the correct relative $x$ position compared to robots 1 to $k-1$.
Finally, all robots are moved right until robot $k$ is in the desired target position.

\subsection{Controlling Covariance Using Wall Friction}\label{subsec:ClosedLoopCovarianceControl}
Assume an open workspace with infinite boundary friction. Goal variances and covariance are  $(\sigma_{goalx}^2,\sigma_{goaly}^2, \sigma_{goalxy})$ and mean, variances and covariance of the swarm  are $( \bar{x},\bar{y},\sigma_x^2,\sigma_y^2, \sigma_{xy})$. For our experiments, $c_1 = 0.1$.
%$\sigma_x^2$ decreases when the swarm is pushed into vertical walls, and $\sigma_y^2$ decreases when the swarm is pushed into horizontal walls. Alg.~\ref{alg:CovCont} procedes as follows:
\begin{enumerate}
\item swarm is pushed into the left wall until $\sigma_x^2< c_1\sigma_{goalx}^2$.  
\item swarm's mean position is moved to the center of the workspace
\item swarm is pushed into the bottom wall until $\sigma_y^2 \le \sigma_{goaly}^2$. 
\item if $\sigma_{goalxy}>0$ swarm slides right until $\sigma_{xy} \ge \sigma_{goalxy}$ \\
else swarm slides left until $\sigma_{xy} \le \sigma_{goalxy}$ 
\item  swarm's mean position is moved to the center of the workspace
\end{enumerate}

>>>>

The region of integration $R$ is the polygon containing the swarm. If $A<1/2$ and the force angle is $\beta$, the mean when the swarm is in the lower-left corner is:
\begin{align}\label{eq:meanInSquareWorkspaceLL}
\bar{x}(A,\beta) &= \frac{\int_0^{\sqrt{2} \sqrt{-A \tan (\beta )}} \left(\int_0^{\sqrt{2} \sqrt{-A \cot (\beta )}+x \cot (\beta )} x \, dy\right) \, dx}{A} \nonumber \\
	&=\frac{ \sqrt{2}}{3} \sqrt{A \tan (\beta )}\\
\bar{y}(A,\beta) &= \frac{\int_0^{\sqrt{2} \sqrt{-A \tan (\beta )}} \left(\int_0^{\sqrt{2} \sqrt{-A \cot (\beta )}+x \cot (\beta )} y \, dy\right) \, dx}{A} \nonumber\\
	&=\frac{\sqrt{2}}{3}  \sqrt{A \cot (\beta )}
\end{align}
The full equations are included in the appendix~\citep{2016arXiv160901830S}, and are summarized in Fig.~\ref{fig:SquareFill}. A few highlights are that the correlation is maximized $\pm$1/2 when the swarm is in a triangular shape. The covariance of a triangle is always $\pm(A/18)$. Variance is minimized in the direction of $\beta$ and maximized orthogonal to $\beta$ when the swarm is in a rectangular shape. The range of mean positions are maximized when $A$ is small.

@@

\subsection{Using Boundaries: Friction and Boundary Layers}\label{subsec:WallFriction}
Global inputs move a swarm uniformly.  
Shape control requires breaking this uniform symmetry.  
A swarm inside an axis-aligned rectangular workspace can reduce variance normal to a wall by simply pushing the swarm into the boundary. 
Controlling covariance by pushing the swarm into a boundary requires changes to the boundary.  
An obstacle in the lower-right corner is enough to generate positive covariance, but generating both positive and negative covariance requires additional obstacles.  
Requiring a special obstacle configuration also makes shape control dependent on the local environment. 
Instead of pushing our robots directly into a wall, this paper examines an oblique approach, by using boundaries that generate friction with the robots.  These frictional forces are  sufficient to break the symmetry caused by uniform inputs.  Robots touching a wall have a friction force that opposes movement along the boundary.  
This causes robots along the boundary to move more slowly than robots in free-space.

Let the control input be a vector force $\vec{F}$ with magnitude $F$ and orientation $\theta$ with respect to a line perpendicular to and into the nearest boundary. $N$ is the normal or perpendicular force between the robot and the boundary. The force of friction $F_f$ is nonzero if the robot is in contact with the boundary and  $|\theta| < \pi/2$. The resulting net force on the robot, $F_{\text{\emph{forward}}}$, is aligned with the wall and given by
\begin{align}
F_{\text{\emph{forward}}} &=  F \sin(\theta) - F_f  \nonumber \\
\text{where }  F_f &= \begin{cases}  \mu_f N, &  \mu_f N < F \sin(\theta)  \label{eq:frictionmodel}  \\
F \sin(\theta), & \text{else} \end{cases} \\ %\sign(F \sin(\theta) ) \cdot  \max(0, | F sin\theta |- |F_f|)
\text{and } N &= F \cos(\theta) \nonumber
\end{align}
 Fig.~\ref{fig:friction} shows the resultant forces on two robots when one is touching a wall. 
Though each receives the same inputs,  they experience different net forces.
  For ease of analysis, the following algorithms assume $\mu_f$ is infinite and robots touching the wall are prevented from sliding along the wall.
This means that if one robot is touching the wall and another robot is free, the touching robot will not move when the control input is parallel or into the wall. 
There are many alternate models of friction that also break control symmetry. Fig.~\ref{fig:friction}c shows fluid flow along a boundary.  Fluid in the free-flow region moves uniformly, but flow decreases to zero in the boundary layer.  
\begin{align}
%u(y) = u_0 \left(1- \frac{(y-h)^2}{h^2}\right) = u_0 \frac{y}{h} \left(2- \frac{y}{h}\right) \label{eq:boundarylayerflow} THiS EQUATION IS TERRIBLE! PLOT IT.
F_{forward}(y) &= F - F_f\begin{cases}  \left( \frac{y-h}{h} \right)^2  , &  y<h \label{eq:boundarylayerflow} \\
0, & \text{else} \end{cases}
\end{align}

The next section shows how a system with friction model \eqref{eq:frictionmodel} and two orthogonal walls can arbitrarily position two robots. 
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.8\columnwidth]{friction.pdf} 
\vspace{-1em}
\caption{(a,b) Wall friction reduces the force for going forward $F_{\text{\emph{forward}}}$ on a robot near a wall, but not for a free robot. (c) velocity of a fluid reduces to zero at the boundary.
\label{fig:friction}
}
\end{center}
\end{figure}

@@

Our algorithm solves the positioning problem in three steps: 
First, $|\Delta r_x - \Delta e_x |$ is reduced to zero while  $\Delta r_y$ is kept constant in Alg.~\ref{alg:XControl}. 
Second, $|\Delta r_y - \Delta e_y |$ is reduced to zero while  $\Delta r_x$ is kept constant. % in Alg.~\ref{alg:YControl}. 
Third, the robots, now correctly positioned relative to each other, are moved to their goal locations.

In the worse case, adjusting both $\Delta r_x$ and $\Delta r_y$ requires two iterations.   Two iterations of Alg. \ref{alg:XControl} are only required if $|\Delta e_x - \Delta s_x|>L$. 
Similarly,  two iterations %of Alg. \ref{alg:YControl} 
are only required if $|\Delta e_y - \Delta s_y|>L$.

@@

\begin{algorithm}
\caption{WallFrictionArrange2Robots($s_1,s_2,e_1,e_2,L$)}\label{alg:PosControl2Robots}
\begin{algorithmic}[1]
\Require starting $(s_1,s_2)$ and ending $(e_1,e_2)$ positions of  two robots. 
$(0,0)$ is bottom corner, $s_1$ is rightmost robot, 
 $L$ is length of the walls. 
 Current robot positions are $(r_1,r_2)$.

\State  GenerateDesired$x$-spacing($s_1,s_2,e_1,e_2,L$)
\State  GenerateDesired$y$-spacing($r_1,r_2,e_1,e_2,L$)
\State Move robots $e_1 - r_1$

@@

\begin{algorithm}
\caption{GenerateDesired$x$-spacing($s_1,s_2,e_1,e_2,L$)}\label{alg:XControl}
\begin{algorithmic}[1]
\Require knowledge of starting $(s_1,s_2)$ and ending $(e_1,e_2)$ positions of  two robots. 
$(0,0)$ is bottom corner, $s_1$ is topmost robot, 
 $L$ is length of the walls. Current robot positions are $(r_1,r_2)$.
\Ensure   $ r_{1y} - r_{2y}  \equiv s_{1y} - s_{2y} $   %$\Delta y(t) \equiv \Delta y(0)$ 
\State $\epsilon \gets $ small number
\State $ \Delta s_x  \gets s_{1x} - s_{2x} $
\State $ \Delta e_x \gets e_{1x} - e_{2x} $
%\State $ r_1 \gets s_1$, $ r_2 \gets s_2$
\If {$\Delta e_x < 0 $ }
\State move$( L-\epsilon-\max( r_{1x},r_{2x}) ,0)   $ \Comment{To right wall}
\Else 
\State  move$ ( \epsilon-\min( r_{1x},r_{2x}),0 )    $ \Comment{Move to left wall}
\EndIf
\State move$(0, -\min( r_{1y},r_{2y} ))$ \Comment{Move to bottom}
%\State $ r_1 \gets r_1+m$, $ r_2 \gets r_2+m$ \Comment{Apply move}
\If {$\Delta e_x - (r_{1x} - r_{2x} ) > 0 $}
\State move$(\min(|\Delta e_x - \Delta s_x |, L- r_{1x}), 0)$  \Comment{Move right}
\Else
\State move$(-\min(|\Delta e_x - \Delta s_x |, r_{1x}), 0)$\Comment{Move left}
\EndIf 
\State move$(0, \epsilon)$ \Comment{Move up}
%\State $ r_1 \gets r_1+m$, $ r_2 \gets r_2+m$ \Comment{Apply move}
%\State $\Delta r_x = r_{1x} - r_{2x}$
\If {$r_{1x} - r_{2x} \equiv \Delta e_x$} 
%\State   $ m \gets (e_{1x}-r_{1x}, e_{1y}-r_{1y})$
%\State $ r_1 \gets r_1+m$, $ r_2 \gets r_2+m$ \Comment{Apply move}
\State  \Return $(r_1,r_2)$
\Else   
\State \Return GenerateDesired$x$-spacing($r_1,r_2,e_1,e_2,L$)
\EndIf
\end{algorithmic}
\end{algorithm}

@@

\begin{figure}
\begin{center}
	\includegraphics[width=1.0\columnwidth]{multirobotSliderHardware.pdf}
\end{center}
\vspace{-1em}
\caption{\label{fig:construction2d}
Illustration of Alg.\ \ref{alg:PosControlNRobots}, $n$ robot position control  using wall friction.
}
\end{figure}

Assume an open workspace with four axis-aligned walls with infinite friction.
The axis-aligned build zone of dimension $(w_b, h_b)$ containing the final configuration of $n$ robots must be disjoint from the axis-aligned staging zone of dimension $(w_s, h_s)$  containing the starting configuration of $n$ robots. Without loss of generality, assume the build zone  is above the staging zone. 
Furthermore, there must be at least $\epsilon$ space above the build zone, $\epsilon$ below the staging zone, and $\epsilon + 2r$ to the left of the build and staging zone, where $r$ is the radius of a robot.  The minimum workspace is then $(\epsilon + 2r + \max(w_b,w_s), 2\epsilon + h_s,h_b)$.

@@

\begin{figure}
\begin{center}
	\includegraphics[width=.9\columnwidth]{driftmove.pdf}
\end{center}
\vspace{-1em}
\caption{\label{fig:driftmove}
A  $\operatorname{DriftMove}(\alpha, \beta, \epsilon)$ to the right repeats a triangular movement sequence $\{ (\beta/2,-\epsilon),(\beta/2,\epsilon),(-\alpha,0)\}$. Robot $A$ touching a top wall moves right $\beta$ units, while robots not touching the top move right $\beta-\alpha$.}
\end{figure}

Let $(0,0)$ be the lower left corner of the workspace, $p_k$ the $x,y$ position of the $k$th robot, and $f_k$ the final $x,y$ position of the $k$th robot. Label the robots in the staging zone from left-to-right and bottom-to-top, and the $f_k$ configurations top-to-bottom and right-to-left as shown in Fig.~\ref{fig:construction2d}.

\begin{algorithm}
\caption{PositionControl$n$RobotsUsingWallFriction($k$)}\label{alg:PosControlNRobots}
\begin{algorithmic}[1]
\State move( $-\epsilon, r-p_{ky}$) % move  away from right wall and down till robot k touches bottom

\While{ $p_{kx} > r$} 
\State $\operatorname{DriftMove}(\epsilon, \min(p_{kx} - r,\epsilon), \epsilon)$ left   %drift move left until kth robot touches left wall
\EndWhile

\State $m \gets \operatorname{ceil}(\frac{f_{ky}-r}{\epsilon})$
\State $\beta \gets \frac{f_{ky}-r}{m}$
\State $\alpha \gets \beta - \frac{r - p_{ky}-\epsilon}{m}$
\For{ $m$ iterations}
\State $\operatorname{DriftMove}(\alpha, \beta, \epsilon)$ up   %move kth robot to f_{ky} and leave the rest in position.
\EndFor

\State move($r+\epsilon-f_{kx}, 0$)  % move the group to the left until k is in the correct relative x position
\State move($f_{kx}-r, 0$)

@@

Alg. \ref{alg:PosControlNRobots} proceeds as follows:  
First, the robots are moved left away from the right wall, and down so robot $k$ touches the bottom wall.
Second, a set of $\operatorname{DriftMove}$s are executed that move robot $k$ to the left wall with no net movement of the other robots.
Third, a set of $\operatorname{DriftMove}$s are executed that  move robot $k$ to its target height and return the other robots to their initial heights. 
Fourth, all robots except robot $k$ are pushed left until robot $k$ is in the correct relative $x$ position compared to robots 1 to $k-1$.
Finally, all robots are moved right until robot $k$ is in the desired target position. Running time is $O(n(w+h))$.

\subsection{Controlling Covariance Using Wall Friction}\label{subsec:ClosedLoopCovarianceControl}
Assume an open workspace with infinite boundary friction. Goal variances and covariance are  $(\sigma_{goalx}^2,\sigma_{goaly}^2, \sigma_{goalxy})$ and mean, variances and covariance of the swarm  are $( \bar{x},\bar{y},\sigma_x^2,\sigma_y^2, \sigma_{xy})$. 
%$\sigma_x^2$ decreases when the swarm is pushed into vertical walls, and $\sigma_y^2$ decreases when the swarm is pushed into horizontal walls. Alg.~\ref{alg:CovCont} procedes as follows:
\begin{enumerate}
\item swarm pushed into the left wall until $\sigma_x^2 \le c_1\sigma_{goalx}^2$.  
\item swarm's mean position is moved to the center of the workspace
\item swarm pushed into the bottom wall until $\sigma_y^2 \le  \sigma_{goaly}^2$. 
\item if $\sigma_{goalxy}>0$ swarm slides right until $\sigma_{xy} \ge \sigma_{goalxy}$ \\
else swarm slides left until $\sigma_{xy} \le \sigma_{goalxy}$ 
\item  swarm's mean position is moved to the center of the workspace
\end{enumerate}
The constant $c_1$ in step 1 should be less than 1 to ensure $\sigma_x^2$ stays within tolerance during the subsequent moves.
